\section{Additional Event-Study and Price-Border Checks}
\label{sec:appendix_event_study}

This appendix reports sample support for the redistricting designs used in the paper and cross-sectional border-RD checks for rents and home sales. The price-border checks use the updated external amenity controls: distance to the nearest school, official Chicago Park District park-boundary polygon, major street, and Lake Michigan. The rental specifications also control for distance to the nearest CTA stop.

\subsection{Redistricting Sample Support}

Table~\ref{tab:redistricting_sample_support} reports treatment and control group counts for the 2015 permit and home-sales redistricting designs.

\input{../tasks/event_study_exact_balance/output/redistricting_sample_support_summary_1000ft.tex}

\subsection{Rental Border RD}

Table~\ref{tab:rental_rd_amenity_attenuation} reports the 500ft RentHub rent RD with no controls, hedonic controls, and hedonic plus amenity controls. The amenity controls do not attenuate the stricter-side rent jump.

\begin{table}[H]
    \caption{RentHub RD: Hedonic and Amenity Controls}
    \label{tab:rental_rd_amenity_attenuation}

    \input{../tasks/rental_rd_characteristics/output/rental_rd_rent_attenuation_bw500.tex}
\end{table}

Table~\ref{tab:rental_rd_amenity_balance} checks balance on the external amenity controls in the same rental RD samples.

\begin{table}[H]
    \caption{RentHub RD: External Amenity Balance}
    \label{tab:rental_rd_amenity_balance}

    \input{../tasks/rental_rd_characteristics/output/rental_rd_external_amenity_balance_bw500.tex}
\end{table}

\subsection{Home-Sales Border RD}

Tables~\ref{tab:sales_rd_amenity_strictness} and~\ref{tab:sales_rd_amenity_binary} report analogous 500ft home-sales border regressions. The first table uses the continuous aldermanic stringency score, while the second uses a binary stricter-side indicator. Both specifications include segment-by-quarter fixed effects and cluster standard errors by ward pair.

\begin{table}[H]
    \caption{Home-Sales Border RD: Stringency Score and Amenity Controls}
    \label{tab:sales_rd_amenity_strictness}

    \input{../tasks/sales_border_pair_fe/output/fe_table_sales_bw500_year_quarter_amenity_clust_ward_pair.tex}
\end{table}

\begin{table}[H]
    \caption{Home-Sales Border RD: Stricter-Side Indicator and Amenity Controls}
    \label{tab:sales_rd_amenity_binary}

    \input{../tasks/sales_border_pair_fe/output/fe_table_sales_right_bw500_year_quarter_amenity_clust_ward_pair.tex}
\end{table}
