\section{Introduction}
\label{sec:intro}

Housing costs in the United States have risen dramatically, and a large body of research points to land use regulation as a key driver. Restrictive zoning, minimum lot sizes, and density caps limit the supply of housing, raise prices, and distort the allocation of labor across cities \citep{glaeser_why_2005, saiz_geographic_2010, turner_land_2014, hsieh_housing_2019, duranton_urban_2023}. But the formal rules on the books are only part of the story. A growing literature emphasizes that the \textit{process} of navigating the regulatory system, including permitting delays, discretionary review, and uncertainty about approval, may matter as much or more than the rules themselves \citep{mayer_residential_2000, paciorek_supply_2013, molloy_housing_2022, soltas_how_2026}. When developers face long and unpredictable timelines, projects shrink, marginal investments become unviable, and housing supply falls \citep{cunningham_house_2006, bloom_impact_2009, oh_land_2024}. Despite the growing recognition that regulatory process frictions are important, there is relatively little causal evidence on how the discretionary authority of local actors shapes the development environment and housing supply.

This paper studies regulatory discretion in the context of Chicago's ``aldermanic privilege,'' the informal power of each of Chicago's 50 city council members (aldermen) to unilaterally block or approve development in their ward. Chicago has no city charter, so there is no statutory basis for this power. It operates through deference: other aldermen vote with the local member on land use matters, city agencies follow the local alderman's lead, and developers understand that meaningful projects require the alderman's blessing \citep{thale_aldermanic_2005, einstein_neighborhood_2019}. The result is that the de facto regulatory environment changes discretely at ward boundaries. Two parcels on opposite sides of the same street, in the same neighborhood, with the same zoning, can face very different regulatory regimes depending on which alderman they fall under. This setting provides unusually clean spatial variation in regulatory discretion, which I use to study its effects on the development process.

To measure the regulatory environment that each alderman creates, I construct ``aldermanic stringency scores'' from permit processing times for high-discretion permits. These are the permit types that tend to draw community opposition and aldermanic intervention: new construction, renovation, demolition, and similar categories that involve meaningful regulatory review. I residualize processing times against ward characteristics and time effects to strip out variation driven by geography and demand, then apply empirical Bayes shrinkage \citep{kline_systemic_2022, kline_discrimination_2024} to address noise for aldermen with few permits. The resulting scores reveal substantial heterogeneity in regulatory stringency across Chicago's 50 wards that is not explained by location or demographics. This approach is motivated by the finding that permitting frictions are a primary determinant of housing supply elasticity \citep{paciorek_supply_2013, gyourko_local_2021, soltas_how_2026}, and by a simple time-to-build framework in which the effective delay imposed by the regulatory process acts as a wedge on the present value of development \citep{kydland_time_1982, mayer_land_2000}.

I use these scores, which are proxies for the regulatory environment, in two empirical designs. The first is a border discontinuity design comparing the density of new multifamily construction on opposite sides of ward boundaries, within narrow bands where neighborhood characteristics are nearly identical but the alderman changes. Using border-segment fixed effects to examine new residential construction density within each ward border, I find that a one standard deviation increase in aldermanic stringency is associated with roughly 10\% lower multifamily density. 
This effect is present for both floor-area ratio and dwelling units per acre, is concentrated near ward boundaries, passes placebo tests at non-boundary locations, and is robust to a range of specification checks.

The second design is an event study exploiting the 2015 ward redistricting. When ward boundaries are redrawn for population balancing, some census blocks are reassigned to new aldermen while nearby blocks remain under the same representation. Using within-census block variation, I find that blocks moved to more stringent aldermen experience a 9\% decline in high-discretion permit issuance over the following five years. The effect is specific to the types of permits that require aldermanic engagement and is much smaller and statistically insignificant for the low-discretion permits (excluding signs) used as a placebo comparison group. This is direct evidence that the development pipeline responds when the regulatory gatekeeper changes.

Finally, I present suggestive evidence that these supply restrictions induced my more stringent aldermen pass through into housing prices. Using the same redistricting variation, I find that home sale prices rise approximately 1-2\% on blocks reassigned to more stringent aldermen relative to nearby blocks that were not reassigned. The estimates are directionally consistent with the supply story, but they are noisier and harder to separate from demand-side explanations, so I treat them cautiously.

This paper contributes most directly to the literature on how local political actors and institutions shape housing supply. \citet{khan_decentralized_2021} studies the same Chicago setting and finds that aldermen approve smaller rezonings near ward boundaries where benefits spill over to adjacent wards. I study a broader margin: by examining realized new construction and not just statutory changes in allowed density via rezoned parcels, I capture the response of the entire development pipeline to aldermanic frictions, including the projects that proceed by-right at reduced scale precisely to avoid triggering discretionary review. 

More broadly, the paper connects to two strands of the urban economics literature. The first is the work on time-to-build and regulatory process frictions. \citet{paciorek_supply_2013} shows that regulation reduces housing supply primarily through its effect on new construction timelines. \citet{mayer_residential_2000} and \citet{mayer_land_2000} demonstrate that development lags are a first-order determinant of the supply response to demand shocks. \citet{soltas_how_2026} provides new evidence that permitting costs are large enough to materially affect housing affordability. My paper complements this work by providing causal evidence that variation in the discretionary layer of the regulatory process, not just the formal rules, generates economically meaningful differences in housing supply within a single city.

The second strand is the literature on how the fragmentation of decision-making authority leads to undersupply of housing. \citet{fischel_homevoter_2009} argues that homeowners use local political institutions to restrict development and protect property values. Subsequent work has documented how this plays out across municipalities, with each jurisdiction restricting development to a degree that may be locally rational but is collectively costly \citep{hilber_origins_2013, ortalo-magne_political_2014, kulka_how_2022, monarrez_dividing_2023, mast_warding_2024, bordeu_commuting_2025}. My results show that a similar coordination failure operates within a single city when informal authority is fragmented across local representatives, each of whom controls the regulatory environment in their own territory.

The rest of the paper proceeds as follows. Section~\ref{sec:data} describes the institutional setting and data. Section~\ref{sec:model} presents a short time-to-build framework that organizes the empirical predictions. Section~\ref{sec:empirical_results_density} presents the cross-sectional density evidence at ward boundaries. Section~\ref{sec:empirical_results_permits} presents the permit event study. Section~\ref{sec:empirical_results_home_sales} presents the home sales results. Section~\ref{sec:discussion} discusses interpretation and policy implications, and Section~\ref{sec:conclusion} concludes.
