\documentclass[12pt]{article}

% PACKAGES
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage[hyperfootnotes=false]{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{verbatim}
\usepackage{float}
\usepackage{caption}
\usepackage{threeparttable}
\usepackage{enumitem}
\usepackage[nomarkers, nolists, figuresfirst]{endfloat}
\renewcommand{\efloatseparator}{\bigskip}

% SETTINGS
\doublespacing

% Custom command for figure notes
\newcommand{\figurenotes}[1]{\par\vspace{0.5em}\footnotesize\textit{Notes:} #1}

\begin{document}

\begingroup
\renewcommand{\thefootnote}{\fnsymbol{footnote}}
\begin{center}
{\Large Regulatory Discretion and Housing Supply: \\ 
Evidence from Chicago's Aldermanic Privilege}

\vspace{1.5em}
Jacob Herbstman\footnotemark[2]

\vspace{0.5em}
\today
\end{center}

\footnotetext[2]{University of Chicago, Harris School of Public Policy.
Email: \url{jherbstman@uchicago.edu}.
\par \smallskip I thank Eric Zwick, Justin Marlowe, Joshua Gottlieb, Asad Khan, Chris Berry, Ayo George, and seminar participants at the Harris School PhD workshop,
Economics Department Student Applied Micro Lunch, and the Notre Dame Housing Economics Conference for helpful comments and suggestions.}
\endgroup
\setcounter{footnote}{0}

\vspace{1em}

\begin{abstract}
Land use regulation constrains housing supply, but formal zoning rules are only part of the regulatory environment developers face.
I study how informal regulations interact with housing supply in the contex of Chicago's ``aldermanic privilege,'' under which city council members can block, delay, or reshape development in their wards.
I construct alderman-level ``regulatory stringency scores'' from permit processing times and compare new construction projects near the ward boundaries, where the neighborhood characteristics and demand-side factors are similar but the regulatory environment changes sharply.
At these boundaries, I find that moving to a more stringent alderman lowers the density of new construction substantially.
I complement this boundary design with an event study exploiting ward redistricting in 2015 and find that a one-standard-deviation increase in aldermanic stringency reduces high-discretion permit applications by $13\%$ over the next 5 years.
I finally look at home prices and rents near ward boundaries and find suggestive evidence consistent with a supply effect: listed rents are about 3.5\% higher and home sale prices about 1.7\% higher on the more stringent side of ward boundaries.
These results emphasize the importance of the informal regulatory environment in shaping housing outcomes, and offers a rationale for why some attempts at formal regulatory reform have failed to increase housing supply.
\end{abstract}

\clearpage

\input{sections/intro.tex}
\input{sections/data.tex}
\input{sections/model.tex}
\input{sections/empirics_results.tex}
\input{sections/discussion.tex}
\input{sections/conclusion.tex}

\newpage

\bibliographystyle{aea}
\bibliography{sections/aldermanic_privilege}

\newpage

% Process main text floats before appendices
\processdelayedfloats
\clearpage

\appendix

\numberwithin{figure}{section}
\numberwithin{table}{section}
\numberwithin{equation}{section}

% Appendix A
\input{sections/appendix_A_data.tex}
\processdelayedfloats
\clearpage

% Appendix B
\input{sections/appendix_B_permits.tex}
\processdelayedfloats
\clearpage

% Appendix C
\input{sections/appendix_C_alderman_fe.tex}
\processdelayedfloats
\clearpage

% Appendix D
\input{sections/appendix_D_density.tex}
\processdelayedfloats
\clearpage

% Appendix E
\input{sections/appendix_E_price_rd.tex}
\processdelayedfloats
\clearpage

\end{document}
