\section{Conclusion}
\label{sec:conclusion}

This paper studies how informal regulatory discretion over development affects housing supply.
I focus on Chicago's aldermanic privilege, the informal power of each of the city's 50 aldermen to unilaterally block or approve development in their ward, and show that it has measurable effects on development activity in the city.

Using permit processing times for projects where aldermen are most likely to matter, I construct aldermanic stringency scores that isolate the alderman-specific component of regulatory delay after removing variation attributable to ward characteristics.
A boundary comparison shows that new construction is less dense on the more stringent side of ward boundaries, with the largest discontinuities for multifamily projects.
The redistricting event study shows that when blocks are reassigned to more stringent aldermen, issuance of high-discretion permits falls by about 7\%.
I also find suggestive downstream evidence on housing costs: listed rents are about 3\% higher and home sale prices are about 1.5\% higher on the more stringent side of ward boundaries after adjusting for observed unit and amenity differences.
The supply-side designs provide causal evidence that variation in informal discretionary authority, beyond the formal rules on the books, affects housing production within a single city.
The results fit a broader literature showing that permitting frictions and time-to-build are important determinants of supply elasticity \citep{paciorek_supply_2013, soltas_how_2026}.
The policy implication is straightforward: reforms to the structure of land use decision-making in Chicago could materially affect housing production.
But as the institutional evidence in Section~\ref{sec:data} makes clear, aldermen control both the permitting process and the zoning code, so effective reform likely needs to address both channels.
