% !TEX root = ../paper.tex

\section{Introduction}
\label{sec:intro}

Housing costs in the United States have risen sharply, and a large literature points to land use regulation as a central cause.
Restrictive zoning, minimum lot sizes, and density caps limit housing supply, raise prices, and distort the allocation of labor across cities \citep{glaeser_why_2005, saiz_geographic_2010, turner_land_2014, hsieh_housing_2019, duranton_urban_2023}.
Formal rules, however, are only one part of the regulatory environment.
A small but growing body of work argues that the political environment and fragmentation of regulatory authority matter as much as the formal regulations themselves \citep{mayer_land_2000, hilber_origins_2013, khan_decentralized_2021, mast_warding_2024, soltas_how_2026}.
When developers face long and unpredictable timelines, projects shrink, marginal investments become unviable, and housing supply falls \citep{cunningham_house_2006, bloom_impact_2009, oh_land_2024}.
Despite this preliminary evidence, there is little empirical work on how individual politicians use these informal regulatory powers to shape housing supply.

This paper studies this informal regulatory power in the context of Chicago's ``aldermanic privilege,'' the unwritten but influential power of each of Chicago's 50 city council members (aldermen) to unilaterally block or approve development in their ward.
Chicago has no city charter, so there is no statutory basis for this power.
It operates through deference: other aldermen vote with the local member on land use matters, city agencies follow the local alderman's lead, and developers understand that major projects require the alderman's blessing \citep{thale_aldermanic_2005, einstein_neighborhood_2019, khan_decentralized_2021}.
This creates sharp changes in regulatory authority at ward boundaries, even when the neighborhood and zoning are quite similar.
Two parcels on opposite sides of the same street, in the same neighborhood, with the same zoning, can face very different regulatory regimes depending on which alderman they interact with.
I use this sharp change in regulatory authority to study how informal regulations affect housing supply and prices.

To measure the regulatory environment each alderman creates, I construct ``regulatory stringency scores'' from processing times for permits likely to involve substantial aldermanic discretion.
I residualize processing times against ward characteristics and time effects to isolate the alderman-specific component of the variation, then apply empirical Bayes shrinkage \citep{kline_systemic_2022, kline_discrimination_2024} to reduce noise for aldermen with relatively few permits.
The motivation for using processing time variation is simple: delay makes projects less valueable, and
a developer who expects years of meetings, redesigns, and uncertainty may build a smaller project or be deterred from building at all \citep{kydland_time_1982, mayer_land_2000, paciorek_supply_2013, gyourko_local_2021, soltas_how_2026}.

I then estimate effects on housing supply by comparing new construction density near ward boundaries.
On one side, developers deal with one alderman; across the street, they deal with another, which can result in a completely different regulatory environment.
The neighborhood and statuatory zoning are similar, but the informal regulatory environment changes sharply.
I use a spatial regression discontinuity design to compare new construction projects on either side of ward boundaries, and find that moving to a more stringent alderman substantially reduces the density of new construction, particularly for multifamily projects.
Multifamily FAR is about $23\%$ lower and DUPAC about $27\%$ lower on the more stringent side of the boundary.
A continuous specification based on the score difference between neighboring aldermen gives the same pattern: at a boundary where their scores differ by one standard deviation, multifamily FAR is about $21\%$ lower and DUPAC about $16\%$ lower on the more stringent side.
The all-construction estimates are much smaller.

I then complement this analysis by using the 2015 ward redistricting to test whether permitting changes when a block is reassigned to a different ward, keeping location fixed.
I estimate the stringency scores for this design using only permits filed through 2014 and measure the assigned change using the origin and destination wards' pre-redistricting incumbents.
Using this within-census-block variation, I find that a one-standard-deviation increase in assigned stringency reduces the annual number of high-discretion permits recorded by the city by 9.4\% in the five years following the remap, relative to nearby blocks whose ward assignment does not change.
This result shows that more stringent aldermen reduce not only the scale of completed projects, but also observed permit activity for projects requiring discretionary review.

I finally examine whether housing costs are higher in the places where the supply evidence predicts they should be.
I once again focus on areas near ward boundaries using a border-discontinuity design.
Listed rents are about 1.7\% higher on the more stringent side of a ward border after controlling for listing characteristics and nearby amenities.
The home-sale estimate points to prices about 1.2\% higher on the more stringent side, but is statistically imprecise.
These results are consistent with lower supply on the more stringent side of ward boundaries and suggest that the uncertainty induced by informal regulations and political actors can have direct effects on affordability.

The paper contributes most directly to work on local control over housing supply and how incentives of particular neighborhoods and the aggregate city may not be aligned. 
\citet{khan_decentralized_2021} studies the same Chicago setting and finds that aldermen approve smaller rezonings near ward boundaries where benefits spill over to adjacent wards.
I study a broader margin by examining realized new construction rather than only statutory changes in allowed density through rezoned parcels to measure how the built environment can change in response to the regulatory environment.

The paper also relates to work on permitting, time-to-build, and fragmented land-use authority.
\citet{mayer_land_2000} and \citet{mayer_residential_2000} show that development lags are a first-order determinant of the supply response to demand shocks.
\citet{paciorek_supply_2013} shows that regulation reduces housing supply primarily through its effect on new construction timelines.
\citet{soltas_how_2026} provides within-parcel evidence that permitting costs are large and that developers value lots that have permits ``ready-to-issue (RTI)''.
This paper shows that the informal layer of the regulatory process and the individual decisionmakers involved can impact housing supply as well.

I also connect to the literature on how the fragmentation of decision-making authority leads to undersupply of housing.
\citet{fischel_homevoter_2009} argues that homeowners use local political institutions to restrict development and protect property values.
Subsequent work has documented how this plays out across municipalities, with each jurisdiction restricting development to a degree that may be locally rational but is collectively costly \citep{hilber_origins_2013, ortalo-magne_political_2014, kulka_how_2022, monarrez_dividing_2023, mast_warding_2024, bordeu_commuting_2025}.
My results show that a similar coordination failure operates within a single city when informal authority is fragmented across local representatives, who can create substantial variation in the regulatory environment that changes discretely at political boundaries.

The rest of the paper proceeds as follows.
Section~\ref{sec:data} describes the institutional setting and data.
Section~\ref{sec:model} presents a stylized time-to-build framework that organizes the supply predictions.
Section~\ref{sec:empirical_results} presents the main results.
Section~\ref{sec:discussion} discusses interpretation and policy implications, and Section~\ref{sec:conclusion} concludes.
