\documentclass[12pt]{article}

% PACKAGES
\usepackage[margin=1in]{geometry}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{setspace}
\usepackage{footnote}
\usepackage{verbatim}
\usepackage{float}
\usepackage{caption}
\usepackage{threeparttable}
\usepackage{enumitem}
\usepackage[nomarkers, nolists]{endfloat}
\renewcommand{\efloatseparator}{\bigskip}

% SETTINGS
\doublespacing

% Custom command for figure notes
\newcommand{\figurenotes}[1]{\par\vspace{0.5em}\footnotesize\textit{Notes:} #1}

\begin{document}

\begin{center}
{\Large Regulatory Discretion and Housing Supply: \\ 
Evidence from Chicago's Aldermanic Privilege}

\vspace{1.5em}
Jacob Herbstman$^\dagger$

\vspace{0.5em}
\today
\end{center}

\renewcommand{\thefootnote}{$\dagger$}
\footnotetext{University of Chicago, Harris School of Public Policy. Email: \url{jherbstman@uchicago.edu}. \par \smallskip I thank Eric Zwick, Justin Marlowe, Joshua Gottlieb, Ayo George, and seminar participants at the Harris School PhD workshop,
Economics Department Student Applied Micro Lunch, and the Notre Dame Housing Economics Conference for helpful comments and suggestions.}
\renewcommand{\thefootnote}{\arabic{footnote}}
\setcounter{footnote}{0}

\vspace{1em}

\begin{abstract}
Stringent land use regulations restrict housing supply and are a key driver of the housing affordability crisis in the United States.
Most research focuses on formal regulations such as zoning codes, while the role of informal regulatory discretion is less understood.
I study how discretionary authority over development affects housing supply using Chicago's ``aldermanic privilege''---the informal power of each city council member to unilaterally block or approve development in their ward.
Using permit processing times to construct a proxy for alderman-level regulatory stringency, I document large differences in the regulatory environment across wards that are not explained by geography or demographics.
A border discontinuity design shows that new multifamily construction is roughly 10\% less dense on the more stringent side of ward boundaries, and an event study exploiting ward redistricting finds that blocks reassigned to more stringent aldermen experience a 9\% decline in high-discretion permits over the following five years.
I also present suggestive evidence that these supply restrictions pass through into higher home prices.
The results suggest that informal discretionary authority is a quantitatively important determinant of housing supply, and that reforms targeting the structure of local land use decision-making could meaningfully affect housing affordability.
\end{abstract}

\clearpage

\input{sections/intro.tex}
\input{sections/data.tex}
\input{sections/model.tex}
\input{sections/empirics_density.tex}
\input{sections/empirics_permits.tex}
\input{sections/empirics_home_sales.tex}
\input{sections/discussion.tex}
\input{sections/conclusion.tex}

\newpage

\bibliographystyle{aea}
\bibliography{sections/aldermanic_privilege.bib}

\newpage

% Process main text floats before appendices
\processdelayedfloats
\clearpage

\appendix

\numberwithin{figure}{section}
\numberwithin{table}{section}
\numberwithin{equation}{section}

% Appendix A
\input{sections/appendix_A_data.tex}
\processdelayedfloats
\clearpage

% Appendix B
\input{sections/appendix_B_permits.tex}
\processdelayedfloats
\clearpage

% Appendix C
\input{sections/appendix_C_alderman_fe.tex}
\processdelayedfloats
\clearpage

% Appendix D
\input{sections/appendix_D_density.tex}
\processdelayedfloats
\clearpage

\end{document}
